---
title: "Module membership, intramodular connectivity and screening for intramodular hub genes (threshold 0.90)"
author: "Ruben Sancho, Bruno Contreras Moreira"
output:
  html_document: default
  pdf_document: default
---

```{r knitr setup, include=FALSE,  eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(eval=TRUE, cache=FALSE, message=FALSE, warning=FALSE,
                      comment = "")
```

#https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Simulated-06-RelatingToExt.pdf
#https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Simulated-07-Membership.pdf


```{r load_Data}
# Load the WGCNA package
library(WGCNA);
library(cluster);
library(ggplot2)
options(stringsAsFactors = FALSE);
allowWGCNAThreads();

getwd()
setwd(".")

# Load data and networks
load(file = "datExpr_unsigned.RData");
load(file = "net_D_unsigned.RData")
```

```{r top_hubs}
# Top hub genes for each module:

chooseTopHubInEachModule(
   datExprD, 
   moduleColors_D, 
   omitColors = "grey", 
   power = 6, 
   type = "unsigned")
```

#Representing modules by eigengenes and relating eigengenes to one another
##To  get  a  sense  of  how  related  the  modules  are  one  can  summarize  each  module  by  its  eigengene  (first  principal component).

```{r relating_eigengenes}
#Calculates module eigengenes (1st principal component) of modules in a given single dataset.
datME_D = moduleEigengenes(datExprD, moduleLabels_D, softPower = 6)$eigengenes

# significant digits (signif) to be used.
datME_D_table = signif(cor(datME_D, use="p"), 2)
write.csv(datME_D_table, file = "datME_D_table.csv")
```

# Intramodular connectivity
# Calculating the intramodular connectivity for each gene.  (In network literature, connectivity is often referred to as ”degree”)
# The function intramodularConnectivity computes:
#--> kTotal: the whole network connectivity (adjacency). Connectivity of the each gene based on its r-values to all other genes in the whole network
#--> kWithin: the within module connectivity (intramodular connectivity). Connectivity of the each gene within a single module based on its r-values to all other genes within the same module
#--> kOut=kTotal-kWithin
#--> kDiff=kIn-kOut=2*kIN-kTotal


```{r Intramodular_connectivity}

# Load adj_D computed previously

load(file = "adj_D.RData")

#Calculates intramodular connectivity, i.e., connectivity of nodes to other nodes within the same module.
Alldegrees_D_colors = intramodularConnectivity(adj_D, moduleColors_D)
head(Alldegrees_D_colors)

```

#Generalizing intramodular connectivity for all genes on the array

##The intramodular connectivity measure is only defined for the genes inside a given module.  But in practice it can be very important to measure how connected a given genes is to biologically interesting modules. A module eigengene-based connectivity measure for each gene as the correlation between a the gene expression and the module eigengene. In summary, how is the relationship between a gene and any module.

#output --> We have a module membership value for each gene in each module.

```{r  intramodular_connectivity_all_genes}

## KME=MM

datKME_D = abs(signedKME(datExprD, MEs_D, outputColumnName="MM."))

# Display the first few rows of the data frame

head(datKME_D)
write.csv(datKME_D, file = "MM_D.csv")
```

#same as previous but changing numbers of modules (1, 2, 3,...) to colors (black, red,...)

```{r  intramodular_connectivity_colors}

# Calculates module eigengenes (1st principal component) of modules in a given single dataset.

datME_D_colors = moduleEigengenes(datExprD,moduleColors_D, softPower = 6)$eigengenes


signif(cor(datME_D_colors, use="p"), 2)
datKME_D_colors = abs(signedKME(datExprD, datME_D_colors, outputColumnName="MM."))
head(datKME_D_colors)
```

#Finding genes with high intramodular connectivity in interesting modules

```{r  HUB_high_intramodular_connectivity_genes}
## module membership measures --> MM


# 1 (turquoise)
FilterGenes_1 = abs(datKME_D$MM.1)>.90
table(FilterGenes_1)
dimnames(data.frame(datExprD))[[2]][FilterGenes_1]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_1], file = "./hub_1.txt",sep="\t",row.names=F,col.names=F,quote=F )


# 2 (blue)
FilterGenes_2 = abs(datKME_D$MM.2)>.90
table(FilterGenes_2)
dimnames(data.frame(datExprD))[[2]][FilterGenes_2]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_2], file = "./hub_2.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 3 (brown)
FilterGenes_3 = abs(datKME_D$MM.3)>.90
table(FilterGenes_3)
dimnames(data.frame(datExprD))[[2]][FilterGenes_3]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_3], file = "./hub_3.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 4 (yellow)
FilterGenes_4 = abs(datKME_D$MM.4)>.90
table(FilterGenes_4)
dimnames(data.frame(datExprD))[[2]][FilterGenes_4]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_4], file = "./hub_4.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 5 (green)
FilterGenes_5 = abs(datKME_D$MM.5)>.90
table(FilterGenes_5)
dimnames(data.frame(datExprD))[[2]][FilterGenes_5]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_5], file = "./hub_5.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 6 (red)
FilterGenes_6 = abs(datKME_D$MM.6)>.90
table(FilterGenes_6)
dimnames(data.frame(datExprD))[[2]][FilterGenes_6]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_6], file = "./hub_6.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 7 (black)
FilterGenes_7 = abs(datKME_D$MM.7)>.90
table(FilterGenes_7)
dimnames(data.frame(datExprD))[[2]][FilterGenes_7]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_7], file = "./hub_7.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 8 (pink)
FilterGenes_8 = abs(datKME_D$MM.8)>.90
table(FilterGenes_8)
dimnames(data.frame(datExprD))[[2]][FilterGenes_8]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_8], file = "./hub_8.txt",sep="\t",row.names=F,col.names=F,quote=F )

# 9 (magenta)
FilterGenes_9 = abs(datKME_D$MM.9)>.90
table(FilterGenes_9)
dimnames(data.frame(datExprD))[[2]][FilterGenes_9]
write.table (dimnames(data.frame(datExprD))[[2]][FilterGenes_9], file = "./hub_9.txt",sep="\t",row.names=F,col.names=F,quote=F )

```



```{r  MM_&_intramodular_connectivity_relationship}

sizeGrWindow(12,12)
png("MM_kIM_cor_D.png", width = 3600, height = 4200, units = "px", pointsize = 32, bg = "white");
par(mfrow=c(10,4), mar=c(5.1,5.1,5.1,5.1))

##power=6

# 1 (turquoise)
which.color= "turquoise";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "1_turquoise",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 2 (blue)
which.color="blue";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "2_blue",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 3 (brown)
which.color="brown";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "3_brown",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 4 (yellow)
which.color="yellow";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "4_yellow",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 5 (green)
which.color="green";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "5_green",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 6 (red)
which.color="red";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "6_red",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 7 (black)
which.color="black";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "7_black",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 8 (pink)
which.color="pink";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "8_pink",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6");

# 9 (magenta)
which.color="magenta";
restrictGenes=moduleColors_D==which.color
verboseScatterplot(Alldegrees_D_colors$kWithin[ restrictGenes],
(datKME_D_colors[restrictGenes, paste("MM.", which.color, sep="")])^6,
col="black", main = "9_magenta",
xlab="Intramodular Connectivity",
ylab="(Module Membership)^6")

```
